Scaling up the hyperparameter grid for Naomi

### 2023-07-11 src/dev_scale-grid/scale-grid.Rmd
Work towards arbitrary orthogonal partial basis AGHQ #58

  **Task** We would like to begin to integrate over the hyperparameters using multiple nodes. The problem is that there are >20 hyperparameters, and so using any type of dense grid is not practical. In this notebook we investigate approaches to using a smaller number of grid points to cover the space, including 1) allocating points in proporiton to posterior standard deviation 2) using different varieties of principal components analysis.
editor_options: 
  markdown: 
    wrap: sentence
---

@@

@@

Remove any nodes which have small enough weight.

>>>>

  **Task** We would like to begin to integrate over the hyperparameters using multiple nodes. The problem is that there are >20 hyperparameters, and so using any type of dense grid is not practical. In this notebook we investigate approaches to using a smaller number of grid points to cover the space, including 1) allocating points in proporiton to posterior standard deviation 2) using different varieties of principal components analysis.
editor_options: 
  markdown: 
    wrap: sentence
bibliography: citations.bib
---

@@

## Creating quadrature grids with different coordinate systems

Want to use a particular coordinate system, but still the covariance matrix scaling as in adaptive quadrature.

```{r}
m <- c(1, 1.5)
C <- matrix(c(2, 1, 1, 1), ncol = 2)

obj <- function(theta) {
  mvtnorm::dmvnorm(theta, mean = m, sigma = C, log = TRUE)
}

ff <- list(
  fn = obj,
  gr = function(theta) numDeriv::grad(obj, theta),
  he = function(theta) numDeriv::hessian(obj, theta)
)

grid <- expand.grid(
  theta1 = seq(-6, 6, length.out = 600),
  theta2 = seq(-6, 6, length.out = 600)
)

ground_truth <- cbind(grid, pdf = exp(obj(grid)))

plot <- ggplot(ground_truth, aes(x = theta1, y = theta2, z = pdf)) +
  geom_contour(col = multi.utils::cbpalette()[1]) +
  coord_fixed(xlim = c(-6, 6), ylim = c(-6, 6), ratio = 1) +
  labs(x = "", y = "") +
  theme_minimal()

grid_5 <- mvQuad::createNIGrid(dim = 2, type = "GHe", level = 5)
mvQuad::rescale(grid_5, m = m, C = C, dec.type = 1)

pca_grid_5 <- pca_rescale(m = m, C = C, s = 1, k = 5)

plot_points <- function(gg) {
  plot +
    geom_point(
      data = mvQuad::getNodes(gg) %>%
        as.data.frame() %>%
        mutate(weights = mvQuad::getWeights(gg)),
      aes(x = V1, y = V2, size = weights),
      alpha = 0.8,
      col = multi.utils::cbpalette()[2],
      inherit.aes = FALSE
    ) +
    scale_size_continuous(range = c(1, 2)) +
    labs(x = "", y = "", size = "Weight") +
    theme_minimal()
}

plot_points(grid_5) + plot_points(pca_grid_5) + plot_annotation(tag_levels = "A")
```

Try to create a grid similar to A above but using the second principal component rather than the first:

```{r}
eigenC <- eigen(C)
E <- eigenC$vectors
lambda <- eigenC$values

gg_s <- mvQuad::createNIGrid(dim = 1, type = "GHe", level = 5)
nodes_out <- t(E[, 2] %*% diag(sqrt(lambda[2]), ncol = 1) %*% t(mvQuad::getNodes(gg_s)))
for(j in 1:2) nodes_out[, j] <- nodes_out[, j] + m[j] # Adaption
weights_out <- mvQuad::getWeights(gg_s) * as.numeric(mvQuad::getWeights(mvQuad::createNIGrid(dim = 1, type = "GHe", level = 1)))
weights_out <- det(chol(C)) * weights_out # Adaption

custom_grid_5 <- mvQuad::createNIGrid(dim = 2, type = "GHe", level = 1)
custom_grid_5$level <- rep(NA, times = 2)
custom_grid_5$ndConstruction <- "custom"
custom_grid_5$nodes <- nodes_out
custom_grid_5$weights <- weights_out

plot_points(grid_5) / plot_points(pca_grid_5) / plot_points(custom_grid_5) + plot_annotation(tag_levels = "A")

ggsave("pca-demo.png", h = 7, w = 6.25)
```

We know that $C = E \Lambda E^\top = \sum \lambda_i \mathbf{e}_i \mathbf{e}_i^\top$

```{r}
lambda[1] * E[, 1] %*% t(E[, 1]) + lambda[2] * E[, 2] %*% t(E[, 2])
```

We also know that any vector can be written in terms of the eigendecomposition.
For example, the vector `c(1, 1)` can be written as:

```{r}
a <- c(1, 1) %*% E

a[1] * E[, 1] + a[2] * E[, 2]
```

# Testing prerotation

See @jackel2005note.

```{r}
R45 <- matrix(
  c(cos(pi / 2), sin(pi / 2), -sin(pi / 2), cos(pi / 2)), ncol = 2, nrow = 2
)
```

@@

Remove any nodes which have small enough weight [@jackel2005note].
